Identification of protoneutron star g-modes in gravitational-wave data

### 2020-06-17 introduction.tex
minor changes

\textbf{CCSN Simulation and GW emission}\\
The life of massive stars ($8 {\rm M}_\odot-100{\rm M}_\odot$) ends with the collapse of their iron core under their own gravity, leading the formation of a neutron star or a black hole (BH), followed (typically but not necessarily in the BH case)  by the explosion of the star as a supernova. Core-collapse supernova (CCSN) explosions are one most important sources of gravitational-waves (GW) that have not yet been detected by current ground-based observatories. This is because even the most common type of CCSN, the neutrino-driven explosion, have a rate about three per century \cite{Gossan:2016} within our galaxy. The other main type of explosion, those produced by the magneto-rotational mechanism, can produce a more powerful signal and can be detected at distances up to $\sim 5$ Mpc \cite{Gossan:2016}. However, the rate of events of this kind is much lower than the one for the neutrino driven mechanism $\sim 10^{-4} \rm{yr}^{-1}$, which represents less than $1 \%$ of all CCSNe.
Despite all this, collapsing stars produces a complex GW signal which could provide significant clues about the physical processes that occur in the moments after the explosion.

The computational modelling of the core-collapse escenario is challenging due the large number of physical processes involved whose role in the escenario is not completely understood. For instance, there are uncertainties in the stellar evolution of massive stars or in the nuclear and weak interactions necessary for the equation of state (EoS) or the neutrino interactions.  These large number of physical ingredients in addition to the necessary accuracy of the modelling of complex multidimensional interactions requires large computational resources. One simulation of one single progenitor in 3D with accurate neutrino transport and realistic equation of state (EoS) can take several months of intense calculations on a scientific supercomputer facility.

\textbf{Mode analysis and relations}\\
\textbf{Martin s20 simulation and code description}\\
\textbf{This paper method description and previous works}\\
\textbf{Previous work in identifying modes from spectrogram}\\
\textbf{Paper organization}\\

>>>>

\textbf{CCSN Simulation and GW emission}\\
The life of massive stars ($8 {\rm M}_\odot-100{\rm M}_\odot$) ends with the collapse of their iron core under their own gravity, leading to the formation of a neutron star or a black hole (BH), followed (typically but not necessarily in the BH case)  by the explosion of the star as a supernova. Core-collapse supernova (CCSN) explosions are one of the most important sources of gravitational-waves (GW) that have not yet been detected by current ground-based observatories. This is because even the most common type of CCSN, the neutrino-driven explosion supernova, have a rate of about three per century \cite{Gossan:2016} within our galaxy. The other main type of explosion, those produced by the magneto-rotational mechanism, can produce a more powerful signal and can be detected at distances up to $\sim 5$ Mpc \cite{Gossan:2016}. However, the rate of events of this kind is much lower than the one for the neutrino driven mechanism $\sim 10^{-4} \rm{yr}^{-1}$, which represents less than $1 \%$ of all CCSNe.
Despite all this, collapsing stars produces a complex GW signal which could provide significant clues about the physical processes that occur in the moments after the explosion.

The computational modelling of the core-collapse phenomena is challenging due the large number of physical processes involved whose role is not completely understood. There are uncertainties in the stellar evolution of massive stars or in the nuclear and weak force interactions necessary for the equation of state (EoS) or the neutrino interactions.  These large number of physical ingredients in addition to the necessary accuracy of the modelling of complex multidimensional interactions requires large computational resources. One simulation of one single progenitor in 3D with accurate neutrino transport and realistic equation of state (EoS) can take several months of intense calculations on a scientific supercomputer facility.

\textbf{PNS oscillation modes}\\
\textbf{Asteroseismology and universal relationships}\\
\textbf{Introduce the topic of the paper: method to extract from the GW data the Mass and the radius of the PNS as function of time using the universal relations.}
\textbf{Previous work in identifying modes from spectrogram}\\
\textbf{Paper organization}\\

### 2020-06-17 methods.tex
minor changes

\section{Methods description}
In this section, we outline a strategy for estimating the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the proto-neutron star (PNS) and its squared radius \textcolor{red}{(units, solar Mass and km?)} based on
the CCSN gravitational wave observations.
An integral part of this strategy is the universal relationship between the characteristic frequency and the ratio of mass and radius as demonstrated by  \cite{Torres:2019}
on 25 1D simulations with the
{\sc AENUS-ALCAR }code \cite{} and the {\sc CoCoNuT} \cite{} code. Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polynomial regression with
heteroscedastic errors
\begin{equation}
r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i
\end{equation}
where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with  variances $\sigma_i^2$ that increase with $f_i$. The model for frequency-dependent variances is
\begin{equation}
\log \sigma_i=\alpha_0+ \alpha_1 f_i + \alpha_2 f_i^2 + \delta_i
\end{equation}
with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package LMVAR
{\textcolor{red}{add citation} that implements a maximum likelihood approach was used to fit the model. The
best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to the  AIC, i.e. \textcolor{red}{insert correct estimates}
\begin{equation}\label{eq:universal}
r_i=f_i + f_i^3 + \epsilon_i
\end{equation}
 The best-fitting model achieves a coefficient of determination of $R^2=$  \textcolor{red}{insert Rsquared}.
Parameter estimates and their standard errors are given in \textcolor{red}{ add Table xxx}.
The data and fit of the model including 95\% confidence bands are displayed in Figure \ref{Fig:LMVAR}.
\begin{figure}
 \centering
 	\includegraphics[width=0.5\textwidth,height=0.3\textheight]{model.pdf}
 \caption{Data from 25 1D simulations {\sc AENUS-ALCAR } and {\sc CoCoNuT}  code, solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands.} \label{Fig:LMVAR}
\end{figure}

>>>>

\section{Methods description}
In this section, we outline a strategy for estimating the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the proto-neutron star (PNS) and its squared radius \textcolor{red}{(units, solar Mass and km?)} based on
the gravitational wave observations.
An integral part of this strategy is the universal relationship between the characteristic frequency and the ratio of mass and radius as demonstrated by  \cite{Torres:2019}
on 25 1D simulations with the
{\sc AENUS-ALCAR }code \cite{} and the {\sc CoCoNuT} \cite{} code. Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polynomial regression with
heteroscedastic errors
\begin{equation}
r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i
\end{equation}
where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with  variances $\sigma_i^2$ that increase with $f_i$. The model for frequency-dependent variances is
\begin{equation}
\log \sigma_i=\alpha_0+ \alpha_1 f_i + \alpha_2 f_i^2 + \delta_i
\end{equation}
with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package LMVAR
\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model. The
best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to the  AIC, i.e. \textcolor{red}{insert correct estimates}
\begin{equation}\label{eq:universal}
r_i=f_i + f_i^3 + \epsilon_i
\end{equation}
 The best-fitting model achieves a coefficient of determination of $R^2=0.9812$.
Parameter estimates and their standard errors are given in \textcolor{red}{ add Table xxx}.
The data and fit of the model including 95\% confidence bands are displayed in Figure \ref{Fig:LMVAR}.
\begin{figure}
 \centering
 	\includegraphics[width=0.5\textwidth,height=0.3\textheight]{model.pdf}
 \caption{Data from 25 1D simulations {\sc AENUS-ALCAR } and {\sc CoCoNuT}  code, solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands.} \label{Fig:LMVAR}
\end{figure}
